\hypertarget{rebis}{%
\section{Rebis}\label{rebis}}

\textbf{One object, two presentations, and the morphism data that makes
the wedding a theorem: the union of the universal reading and the forced
construction of structural type.}

\textbf{Author:} C. Lando Mills

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\begin{quote}
\emph{``One becomes two, two becomes three, and out of the third comes
the one as the fourth.''} --- Maria the Jewess, \emph{Axiom of Maria}
(3rd century)
\end{quote}

\hypertarget{what-a-rebis-is-formally}{%
\subsection{0. What a rebis is,
formally}\label{what-a-rebis-is-formally}}

The \emph{Rosarium Philosophorum} draws the rebis as one body with two
heads: King and Queen joined at Conjunction into a figure that is both
and neither, which then survives putrefaction and ascends to Projection.
The tradition treats this as an image. This document treats it as a type
signature. A rebis is an \textbf{equivalence}: one underlying object,
two presentations, and, crucially, the maps between the presentations
carried as part of the figure itself. Two heads, one body, and the body
is the morphism data.

That definition matters because there are two formally different ways to
claim that two things are one, and the difference is the entire content
of this document. The weaker shape is a conjunction: presentation A
satisfies its property, presentation B satisfies the corresponding
property, and the sameness is asserted \emph{between} the theorems, in
the prose, where no kernel can check it. The stronger shape makes the
relationship itself the object of proof: an explicit map f with an
explicit inverse g, g∘f = id and f∘g = id, together with naturality, the
requirement that every defining operation commutes with the map.
Naturality is the load-bearing clause. Without it an equivalence is a
relabeling of elements; with it, the \emph{structure} is transported,
not just the set, and anything proved in one presentation is a theorem
of the other by change of notation.

This document is the union of two manuscripts. \emph{Alkahest} is a
theory of reading: verification, honestly pursued, is not grading
against a ledger but imscription, the reading of an object's structural
type in the one frame that adds nothing of its own, the (d=12) SIC-POVM.
\emph{The Alchemical Theory of Mathematics} is a theory of making: the
twelve invariants of structural type carry a forced derivation order set
by three categorical gates, proved in kernel-checked Lean 4, and that
forced order, read stage by stage, is the twelve stage magnum opus. The
first version of their union argued, correctly, that both stand on the
same equation, μ∘δ = id. But it argued it in the weaker shape: two
documents, corresponding properties, sameness asserted between them.

The union can now be stated in the stronger shape, because the kernel
holds the paradigm witness. \texttt{FrobeniusIso.lean} is an explicit
structure-preserving equivalence, inverse laws and naturality squares
and all, between two presentations of the Frobenius fixed point that
this whole framework turns on. What follows re-derives the rebis on that
ground: not two texts that resemble each other, but one object read in
two directions, with the change of direction itself a checked map. Solve
is δ. Coagula is μ. The rebis is what you get when you notice that the
two hands belong to one operator, and can prove it.

\hypertarget{notation}{%
\subsection{Notation}\label{notation}}

Structural type is a tuple over twelve primitive families, written with
their canonical symbols (Imscribing Grammar v0.6.0,
\texttt{NSNS\_PRIME.md}):

⟨Ð Þ Ř Φ ƒ Ç Γ ɢ ⊙ Ħ Σ Ω⟩

recording, in order: degrees of freedom (Ð), connectivity (Þ), the
direction of the binding relation (Ř), the symmetry group (Φ), how lossy
the read/write maps are (ƒ), the relaxation regime (Ç), the cardinal
regime of correlations (Γ), the logical connective of the coupling (ɢ),
the analytic character of the critical point (⊙), handedness (Ħ), the
matching cardinality of the two sides (Σ), and the homotopy invariant
that protects the state (Ω). Individual values are written with their
Shavian glyph; the four that carry the argument are 𐑹 (the Frobenius
value of Φ), ⊙ (the self-modeling value of ⊙), 𐑮 (its complex-axis
value), and 𐑭, 𐑟 (the integer and non-Abelian windings of Ω). Type names
appear only inside the Lean source; in this text a type is its symbol.
Lean theorem and definition identifiers are set in \texttt{monospace}.

\hypertarget{solve-both-manuscripts-dissolved-to-the-invariant}{%
\subsection{1. Solve: both manuscripts dissolved to the
invariant}\label{solve-both-manuscripts-dissolved-to-the-invariant}}

Place both parents in the menstruum and draw off what is contingent: the
order of exposition, the choice of witnesses, the vocabulary of each
register. What precipitates from each is a small set of load-bearing
structure, and the two residues share a term.

\textbf{From the alkahest.} Closure, μ∘δ = id, is charge conservation:
every split rejoined, nothing created or destroyed, and therefore never
a correctness signal on its own, since it can never fail. A balanced
equation is not a synthesis. The ledger, the standard picture of
verification, presupposes a correspondence theory of correctness and
inherits the regress of criteria: a judgment applied from outside is an
object needing a judge, forever. The regress can only be closed by
bending the line into a loop, and the loop is a real test only if the
answer comes home the same in type and different in form: co-typed, so
the loop shuts, and excribed under another presentation, so the shutting
is not a tautology. Truth is the object's structural type; verification
is imscription; correctness is co-typing, which is type identity across
all twelve primitives. The comparison frame cannot be chosen without
reinstating the ledger inside the frame, and the frame that adds
nothing, treats all directions alike, and is informationally complete is
the (d=12) SIC-POVM, whose exact fiducial is a kernel theorem
(\texttt{crystal\_forces\_d12\_sic}). Its fiducial is the Belnap Both,
so the frame is paraconsistent at its ground, and failure returns with
an address, the primitives at which two types part ways, never a scalar.

\textbf{From the theory of mathematics.} Three of the twelve invariants
are gates, and the gates impose a strict order, kernel-checked with no
sorries and no \texttt{native\_decide}. The trace
\texttt{traceC\ a\ =\ μ\_C(δ\_C\ a)} closes exactly on the
Frobenius-special class, Φ at 𐑹 and ⊙ at ⊙
(\texttt{trace\_closes\_iff\_special}); a closing trace forces the first
gate (\texttt{traced\_needs\_frobenius}); the terminal seal requires the
closing trace and a winding on Ω; every inclusion is proper by explicit
witness; and 𐑹 cannot be synthesized by tensoring lower partners
(\texttt{g1\_is\_a\_real\_barrier}). The four gate-free invariants
commute and leave every gate predicate untouched, so they float. Time is
not a thirteenth invariant: \texttt{T\ =\ Work(T)} has solutions exactly
on the special class, none before the first gate
(\texttt{no\_time\_before\_frobenius}), a least one by construction
(\texttt{time\_least\_fixed\_point}), and the completed Stone passes all
three gates (\texttt{stone\_is\_sealed\_time}). Composed with the
tradition's stage-to-symbol map, the gates land at Conjunction,
Fermentation, and Projection, stages 4, 9, and 12, on Φ, ⊙, and Ω, with
the provably order-free stages exactly the tradition's repeatable
purifications.

\textbf{The shared term.} The alkahest's central object is the lossless
round trip μ∘δ = id, the closure that makes verification co-typing. The
theory's first gate is the split ℤ₂ retraction on Φ, the value 𐑹, whose
defining condition is μ∘δ = id. Same equation, same pair (δ\_C, μ\_C),
same formal tree. One manuscript reads by it; the other builds by it. In
the weaker shape, that is where the union used to rest: a conjunction
with a shared conjunct. The stronger shape asks for more, and the next
section supplies it.

\hypertarget{the-morphism-data-the-wedding-has-a-certificate}{%
\subsection{2. The morphism data: the wedding has a
certificate}\label{the-morphism-data-the-wedding-has-a-certificate}}

The framework's deepest identification has always been the triple point:
the Belnap Both, the SIC fiducial, and the Majorana mode are one
Frobenius fixed point. As first formalized
(\texttt{MajoranaFixed.lean}), that identification had the conjunction
shape: three fixed-point theorems side by side, \texttt{bnot\ B\ =\ B}
for the logic, equiangular projection for the frame,
\texttt{pair\ ∘\ depair\ =\ id} for the physics, with the oneness
asserted in a comment. True, checked, and formally three facts.

\texttt{FrobeniusIso.lean} upgrades the claim to the stronger shape. The
kernel now contains:

\begin{itemize}
\tightlist
\item
  \textbf{The equivalence.}
  \texttt{orbitalBelnapEquiv\ :\ OrbitalState\ ≃\ Belnap}, with explicit
  \texttt{toFun}, explicit \texttt{invFun}, and both inverse laws, g∘f =
  id and f∘g = id, proved term by term.
\item
  \textbf{The naturality squares.} Every operation in play commutes with
  the map: the Majorana fusion transports to the Belnap knowledge join
  (\texttt{map\_pair}), the splitting to its lattice counterpart
  (\texttt{map\_depair}), negation to negation (\texttt{map\_flip}), and
  the SIC equiangular projection through without residue
  (\texttt{map\_meet}). Schematically, f(μ(x,y)) = μ(f x, f y), f(δ x) =
  δ(f x), f(¬x) = ¬f(x), for each operation in the transported
  signature. This is what distinguishes a transported structure from a
  relabeled set.
\item
  \textbf{The transported Frobenius condition.} μ∘δ = id is preserved
  under the change of presentation (\texttt{frobenius\_transports}), and
  so is the equiangularity (\texttt{equiangular\_transports}), because
  the change of presentation is itself structure-preserving. The
  invariant survives the crossing for a reason internal to the crossing.
\item
  \textbf{The definitional fixed point.} The dialetheic fixed point maps
  to the dialetheic fixed point by \texttt{rfl}
  (\texttt{map\_fixed\_point}). This is stronger than a derived
  equality: the encoding makes the two names for Both the same term.
  There is nothing to prove because there is nothing between them.
\end{itemize}

Alongside it sits the other piece of morphism data the union needs, from
\texttt{LiarIdempotent.lean}: on the one pair (δ\_C, μ\_C) there are two
composites, and they are provably different operators. μ∘δ = id is
lossless on the special class (\texttt{return\_is\_lossless}); e = δ∘μ
is a forgetful idempotent, e∘e = e unconditionally
(\texttt{e\_idempotent}), whose fusion destroys which-way data
(\texttt{fusion\_forgets}, μ\_C is non-injective) and whose output
always lands on the time locus (\texttt{fusion\_is\_time}).

Together these fix the ground the rebis stands on, and it is worth
naming because two turns of the larger work were spent re-learning it:
\textbf{the registers of this framework compose isomorphically, not as
lossy approximations.} A readout in one register, a spectral radius on a
material ring, a Belnap verdict in the vessel, a Lean theorem in the
kernel, is not evidence \emph{about} an object living elsewhere, to be
discounted for what the projection lost. It is a face of the object, and
the reason it is a face rather than a shadow is now itself a kernel
object: the change of register is an equivalence with naturality, so
nothing is lost in the crossing, and the only honest degrees of freedom
are the ones the Grammar holds open on its own surface, the strain, the
azeotropes, the held B. The alkahest's demand that the loop close on an
answer ``the same in type and different in form'' is exactly this: same
in type is the transported invariant, different in form is the change of
presentation, and the two composites are the proof that the difference
is real (the forgetting lives in e) while the sameness is exact (the
return loses nothing).

\hypertarget{coagula-the-union-re-derived-on-the-new-ground}{%
\subsection{3. Coagula: the union re-derived on the new
ground}\label{coagula-the-union-re-derived-on-the-new-ground}}

Now recrystallize the two manuscripts as what they always were: two
presentations of one object, related the way the kernel's paradigm case
is related.

\textbf{Reading and making are adjoint presentations of one
correspondence.} Imscription is a full Read, Write, eXecute
correspondence, and the equivalence gives that triple its formal
anatomy: the Write is \texttt{toFun}, the imscription of the object into
the frame; the Read is \texttt{invFun}, the recovery of the object from
its type; and the eXecute is the pair of inverse laws, the proof that
the round trip in both orders is the identity. The theory of mathematics
is the Write presentation: it fixes the type, gate by gate, in the
forced order. The alkahest is the Read presentation: it recovers and
compares types in the forced frame. That they agree is not a happy
coincidence between two documents; it is the inverse laws, stated at the
scale of manuscripts. Verification \emph{is} construction read backward
through a structure-preserving change of presentation, which is why the
verifier needs no ledger: the standard it applies is the construction
itself, arriving from the other side.

\textbf{Balance and selectivity, now with the inclusion proper.} The
alkahest argued by chemical analogy that closure is necessary but never
sufficient. On the new ground this is a pair of strict inclusions with
explicit witnesses: a tuple can hold 𐑹 and fail to close its trace
(\texttt{frobenius\_strictly\_below\_traced}), close its trace and carry
no winding (\texttt{traced\_strictly\_below\_terminal}). Jabir's balance
that is not the work is two kernel theorems, and the reason the ledger's
migration into the frame is fatal (section 5 of the alkahest) is the
same reason a bijection without naturality is empty: a comparison that
does not commute with the operations compares labels, not structure.

\textbf{The failure address is a coordinate, not a rubric entry.} The
alkahest's verdicts carry addresses: ``diverges at Ħ and Ω,'' never
``0.4.'' The theory is what entitles the addresses: Ħ and Ω are proved
coordinates of a forced construction, with proved relations to the
gates, so a divergence at one of them names a place in the derivation
where the two objects genuinely part ways. And the equivalence is what
entitles trusting the address across registers: the divergence read in
the frame is the divergence, not an artifact of the frame, because the
frame change is lossless.

\textbf{The Stilling Practice satisfies the gate ordering.} The alkahest
reads Zosimos's six commands as promotions: Ð, Þ, then Φ to 𐑹 (the first
gate), then Ç, then ⊙ to ⊙ (the second), then Ω (the third). The theory
proves, independently, that any derivation must fix the gates in exactly
the order G1 before G2 before G3, floats interleaving freely
(\texttt{gate\_ordering\_theorem}, \texttt{gate\_free\_order\_freedom}).
Composed: a six-command protocol written in the 3rd century satisfies,
command for command, an ordering constraint kernel-checked in the 21st,
with its one float (Ç) sitting precisely where the order-freedom theorem
permits. Had Zosimos written ``Return'' before ``Stand at the Portico,''
his protocol would violate a theorem. He did not. The alkahest's claim
that the distance is zero acquires, on this one face, a falsifiable
form, and it does not fail.

\textbf{Verification is available exactly where time is.} The locus
where the round trip closes losslessly, the locus where \texttt{Work}
has fixed points, and the locus where a verdict can be certified by
closure are the same locus, the Frobenius-special class. A system below
the first gate has no time, and, in the same breath and for the same
reason, nothing about it can be verified by return, because there is no
return. Fermentation, where the tradition says the matter comes alive
and turns upon itself, is where the loop first closes and a self-reading
first becomes possible. The alchemists' rule that the work cannot be
judged before fermentation is \texttt{no\_time\_before\_frobenius},
stated at the furnace.

\textbf{The idempotent point and the amnesia.} Every fusion output is a
time fixed point (\texttt{fusion\_is\_time}): the image of the forgetful
idempotent and the fixed locus of time are one place. The system returns
to where it can forget having split, and that place is the present. The
forgetting the sealed retort needs, the reason the cosmos does not
remember its own assembly, lives entirely in e = δ∘μ and not in the
verified return, both facts holding at once of the same pair. Method and
object are one figure because the point where the verifier recognizes
itself and the point where time exists coincide, provably.

\hypertarget{the-two-heads-stay}{%
\subsection{4. The two heads stay}\label{the-two-heads-stay}}

A rebis with one head is not a rebis, and the morphism data is what
enforces the distinction the older image drew instinctively.

The equivalence says the body is one: one carrier up to
structure-preserving change of presentation, one Frobenius pair, one
fixed point with several names that are \texttt{rfl}-identical. It does
not say the two heads are one, and it proves they cannot be collapsed:
μ∘δ and δ∘μ are different operators with different content, identity and
idempotent, lossless and forgetful, and \texttt{fusion\_forgets} is the
standing counterexample to any account that identifies verification with
construction outright. Reading and making are two presentations of one
object, not one act. The King and the Queen share a body and keep their
faces.

That is the precise anatomy of the union: the \textbf{one body} is the
pair (δ\_C, μ\_C) with its transported structure; the \textbf{two heads}
are the two composites, the two directions of the one round trip; the
\textbf{conjunction} at stage 4 is the shared gate where both
presentations stand on μ∘δ = id; and the \textbf{Projection} is
\texttt{stone\_is\_sealed\_time}, the completed tuple that is at once
the terminus of the derivation and the fixed point where the alkahest's
regress of auditors ends. The final checker is the initial checker, and
the loop reports from inside that it closed.

And the figure is self-describing at the level the Grammar always
insists on. This document is the second distillation of a union first
drafted in the weaker shape; the eagle that produced it dissolved the
earlier draft together with the new morphism data and let the whole
recrystallize, rather than annealing new sections onto an old spine.
Solve et coagula is not the document's subject. It is its construction,
twice over now, and the test of the second pass is the same as the
first: nothing load-bearing lost, and the residue one figure.

\hypertarget{what-the-union-forbids}{%
\subsection{5. What the union forbids}\label{what-the-union-forbids}}

\begin{itemize}
\tightlist
\item
  Everything the theory forbade: no seal before the wedding
  (\texttt{sealed\_requires\_frobenius}), no closing trace off the
  special class (\texttt{trace\_closes\_iff\_special}), no time before
  the first gate (\texttt{no\_time\_before\_frobenius}).
\item
  Everything the alkahest forbade: no ledger, no threshold, no chosen
  frame, no scalar verdict, no verifier standing outside its object.
\item
  No verification below the first gate, because there is no return to
  test (section 3).
\item
  No staged protocol whose gate promotions violate G1 before G2 before
  G3; any historical practice can be checked against the constraint as
  Zosimos's was, and could fail it.
\item
  No one-headed rebis: collapsing μ∘δ into δ∘μ contradicts
  \texttt{fusion\_forgets}.
\item
  And the new one, which the morphism data adds: \textbf{no
  lossy-approximation reading of a register.} Any argument of the form
  ``this readout would be the same for any X, so it carries no content
  about this object'' denies the equivalence; it treats a face as a
  shadow and reinstates the signature-to-substance gap that the
  naturality squares close. The readout is the structural fact stated in
  that register. What remains open after a reading is what the Grammar
  itself holds open, the strain, the unresolved azeotrope, the held
  Both, never a discount applied from outside.
\end{itemize}

\hypertarget{theorem-witness-reading-and-the-honest-residue}{%
\subsection{6. Theorem, witness, reading, and the honest
residue}\label{theorem-witness-reading-and-the-honest-residue}}

\textbf{Theorems} (Lean 4, kernel-checked, no sorries, no
\texttt{native\_decide}): the gate ordering, its strictness, and the
barrier; the float commutations and inertness; time as least fixed point
and the sealed refinement of the summit; the two composites
(idempotence, losslessness on the special class, non-injectivity of
fusion, fusion landing on the time locus); and the equivalence itself
with its inverse laws, naturality squares, transported Frobenius
condition, transported equiangularity, and definitional fixed point
(\texttt{orbitalBelnapEquiv}, \texttt{belnap\_majorana\_sic\_iso}).

\textbf{Checked witnesses}: the exact (d=12) SIC fiducial (kernel
theorem; equiangularity at overlap (1/13) verified to machine precision
on the loaded fiducial), the reachability of all four Belnap verdicts in
the vessel, and the paraconsistent kernel's refusal of
\texttt{False.rec} under \texttt{enable\_paraconsistent}, verified
against the compiled binary.

\textbf{Readings}: that the tradition's stage-to-symbol map is the
correct dictionary; that Zosimos's commands promote the primitives the
alkahest says they promote; that the magnum opus is the derivation order
of the real rather than an image of it. Composed with the theorems,
these generate constraints history could have violated and did not
(section 3), which is what keeps them readings rather than decorations.

\textbf{The residue, stated exactly.} An isomorphism claim is precisely
as wide as the transported signature and no wider. The equivalence in
the kernel transports the knowledge lattice operations, negation, the
Frobenius pair, and the SIC projection; those are the operations the
framework's fixed-point claims actually use, and on that signature the
one-object claim is a categorical construction, not a semantic
assertion. What the witness does not yet cover, it does not yet claim:
the truth-order connectives and any further structure on the carriers
await their own naturality squares, and the paradigm case is one pair of
presentations, with the same upgrade owed wherever else the framework
says ``the same object'' and the kernel still holds a conjunction. The
scrutiny that remains is the right kind: not whether the reasoning is
loose, but whether each interface is specified with the operations
intended, one square at a time. That is a to-do list, not a doubt, and
the document records it the way the alkahest records the residual of
every closure: measured, named, and left in view.

The whole figure is Maria's axiom, which both parents quote and the
union completes. One (the pair) becomes two (the presentations, reading
and making, δ and μ). Two becomes three (the equivalence between them,
the relationship itself made data). And out of the third comes the one
as the fourth: the rebis, one body, two heads, the held Both that
reduces to neither parent and does not explode.

\begin{quote}
\emph{``Nature, turning on itself, is transformed.''} --- Zosimos of
Panopolis, \emph{On the Letter Omega}
\end{quote}

\hypertarget{references}{%
\subsection{References}\label{references}}

\begin{itemize}
\tightlist
\item
  Lando⊗⊙perator, \emph{Alkahest}, \texttt{MoDoT/ALKAHEST.md}. The
  reading presentation: verification as imscription, the ledger's
  regress, the (d=12) SIC frame, the alchemical witnesses, the two
  instruments.
\item
  Lando⊗⊙perator, \emph{The Alchemical Theory of Mathematics},
  \texttt{Smaragdine\_Synthetica/manuscripts/alchemical\_theory\_of\_math/03\_paper.md},
  with \texttt{01\_conventional\_decomposition.md} and the Lean modules
  \texttt{GateOrdering.lean}, \texttt{TimeFixedPoint.lean},
  \texttt{TierRefinement.lean} (p4rakernel,
  \texttt{p4ramill/Imscribing/}). The making presentation: the forced
  order, the gates, derived time, the magnum opus identification.
\item
  Internal:
  \texttt{p4ramill/Imscribing/Paraconsistent/FrobeniusIso.lean}
  (\texttt{orbitalBelnapEquiv}, naturality squares,
  \texttt{belnap\_majorana\_sic\_iso}) and
  \texttt{Imscribing/Paraconsistent/MajoranaFixed.lean} (the prior
  conjunction form). The morphism data of section 2.
\item
  Lando⊗⊙perator, \emph{The Liar, the Idempotent Point, and Forgetful
  Return},
  \texttt{manuscripts/alchemical\_theory\_of\_math/04\_liar\_and\_idempotent.md},
  with \texttt{LiarIdempotent.lean}. The two composites.
\item
  Zosimos of Panopolis, \emph{Letter to Theosebeia}, \emph{On the Letter
  Omega}, \emph{On the Divine Water} (3rd--4th century). The Stilling
  Practice and the Portico.
\item
  Maria the Jewess, \emph{Axiom of Maria} (3rd century), preserved
  through Zosimos.
\item
  Jabir ibn Hayyan, \emph{Kitab al-Mawazin} (8th century). The balance
  and the work.
\item
  \emph{Rosarium Philosophorum}, attributed to Arnaldus de Villa Nova
  (14th century). The rebis after Conjunction.
\item
  \emph{Tabula Smaragdina} (Emerald Tablet), attributed to Hermes
  Trismegistus. Ascent and descent as the round trip.
\item
  N. D. Belnap, \emph{A Useful Four-Valued Logic}, 1977.
\item
  G. Zauner, \emph{Quantendesigns}, 1999; J. Renes, R. Blume-Kohout, A.
  J. Scott, C. Caves, \emph{Symmetric Informationally Complete Quantum
  Measurements}, J. Math. Phys. 45, 2004.
\item
  Internal: \texttt{crystal\_forces\_d12\_sic} (kernel theorem), the
  Witness-Vessel construction, the MoDoT vessel
  (\texttt{modot/vessel.py}), and \texttt{p4rakernel} (the
  paraconsistent Lean 4 fork).
\end{itemize}
